\cabstract{
电梯轿厢是典型的封闭式高频公共微空间，具有反光强、遮挡多、视角近、目标尺度变化快和事件响应时延要求高等特点。针对电动车入梯风险治理中离线检测精度与板端部署表现可能不一致的问题，本文以 HongOU PI（Hi3516DV500）开发板为部署平台，研究电梯轿厢场景中的行人与电动车双目标检测、模型兼容迁移和部署一致性评价。本文旨在解决三个问题：如何将电动车入梯治理抽象为稳定的双目标检测任务，如何保证 YOLO 模型从训练权重到 OM 离线模型及板端应用的语义一致，如何通过可复查实验解释模型迁移后的检测结果、阈值行为和运行时口径。

本文的主要创新在于将 YOLO 双目标检测模型的部署过程契约化和证据化。首先，本文分析电梯轿厢的空间结构、成像特点和安全响应需求，明确 \textit{person} 与 \textit{ebike} 双目标检测任务边界。其次，结合 YOLO 检测器的 Backbone、Neck、Head 和后处理流程，将自训练权重到 ONNX 中间表示、OM 离线模型和 Hi3516DV500 板端推理的过程抽象为输入、类别、输出、后处理和运行五类迁移接口契约。最后，针对模型迁移到板端后可能出现的类别语义、边框解释、阈值行为和运行时口径变化，构建部署一致性评价框架，并以完整验证集、分类别指标、阈值后验重算、NMS 敏感性、后处理消融、分段计时和代表性视频片段作为证据来源。

实验结果表明，在 720 张图像构成的验证集上，边缘设备端总体 precision 为 0.901，recall 为 0.923，F1 为 0.912，mAP@0.5 为 0.910；其中行人类别 mAP@0.5 为 0.869，电动车类别 mAP@0.5 为 0.952。作为训练端参照，Linux 服务器 FP32 自训练模型在同一验证集上的 mAP@0.5 为 0.988，板端 FP16/OM 结果相比训练端下降约 0.078，但仍保持较高检测能力。阈值敏感性后验重算显示，在 0.20--0.40 置信度阈值范围内，总体 F1 保持在 0.915--0.922 区间，电动车召回率保持在 0.953 以上。板端 batch 图像验证链路端到端平均单图处理耗时约为 1120 ms；进一步的分段计时表明，50 张图像 batch 复查中模型执行阶段均值约为 8.2 ms，代表性连续视频处理中模型执行阶段均值约为 8.4 ms。上述结果说明，本文方法在当前验证集与代表性片段范围内表现出较稳定的双目标检测能力，并提供可复查的部署一致性证据。
}

\ckeywords{电梯安全视觉；边缘智能；目标检测；部署一致性；电动车入梯检测}

\eabstract{
Elevator cabins are closed, high-frequency public micro-spaces characterized by strong reflections, frequent occlusions, close viewpoints, rapid scale variation, and strict latency requirements for safety response. To address the potential gap between offline detection accuracy and board-side deployment behavior in e-bike-in-elevator risk governance, this thesis studies pedestrian and e-bike detection in elevator cabins using the HongOU PI (Hi3516DV500) development board as the deployment platform. The goal is to formulate e-bike-in-elevator governance as a stable dual-target detection task, preserve the semantics of a YOLO model from training weights to an OM offline model and board-side application, and explain the deployed model through reproducible detection, threshold, post-processing, and runtime evidence.

The main contribution is to make the YOLO deployment process both contract-based and evidence-based. First, this thesis analyzes the spatial structure, imaging characteristics, and safety requirements of elevator cabins, and defines the dual-target detection task for \textit{person} and \textit{ebike}. Second, the migration from self-trained weights to ONNX, OM offline models, and board-side inference is formulated as five interface contracts: input, category, output, post-processing, and runtime contracts. Finally, to analyze behavior changes after board-side deployment, a deployment consistency framework covering category semantics, bounding-box interpretation, threshold behavior, NMS behavior, post-processing effects, and runtime measurements is constructed using reproducible experimental evidence.

Experimental results show that on a 720-image validation set evaluated on the edge device, the system achieves an overall precision of 0.901, recall of 0.923, F1 of 0.912, and mAP@0.5 of 0.910. The pedestrian class reaches 0.869 mAP@0.5, while the e-bike class reaches 0.952 mAP@0.5. As a training-side reference, the self-trained Linux FP32 model achieves 0.988 mAP@0.5 on the same validation set, so the board-side FP16/OM result shows an mAP@0.5 decrease of about 0.078 while retaining strong detection capability. A posterior threshold recomputation shows that the overall F1 remains within 0.915--0.922 under confidence thresholds from 0.20 to 0.40, and the e-bike recall remains above 0.953. The average end-to-end processing time of the board-side batch validation path is about 1120 ms per image; segmented timing shows that model execution takes about 8.2 ms in a 50-image batch check and about 8.4 ms in a representative video run. These results indicate stable dual-target detection within the current validation set and representative snippets, while providing reproducible evidence for deployment consistency evaluation in elevator edge vision.
}

\ekeywords{Elevator safety vision; Edge intelligence; Object detection; Deployment consistency; E-bike-in-elevator detection}
